Financialisation of Commodity Markets \\\large{Co-movement Behind-the-Scenes}

### 2025-04-04 docs/paper.qmd
Recompile for resubmission

The commodity futures trading commission (CFTC) operates a comprehensive system of collecting information on market participants known as the large trader reporting program (LTRP) where it records and reports to the public position data on market participants who have position levels in excess of a particular market specific threshold^[Current reporting level thresholds available at: [www.ecfr.gov](https://www.ecfr.gov/cgi-bin/retrieveECFR?gp=&SID=970471b8455f4bab7db4110cfde50731&mc=true&r=SECTION&n=se17.1.15_103).]. The commission collects market data and position information daily from clearing members, futures commission merchants (FCMs) and foreign brokers and publishes corresponding summaries weekly in a series of market specific reports. In these reports, individual traders are categorised according to the nature of their trading activity on the basis of self-reported information^[CFTC form 40. Available at: [www.cftc.gov](https://www.cftc.gov/sites/default/files/idc/groups/public/@forms/documents/file/cftcform40.pdf).] that is subject to review by CFTC staff for reasonableness. Position information is then aggregated by category and asset type with each report providing an aggregation for a set of report specific categories and, for each category, a breakdown between futures only positions and positions in futures and options combined.  
In its legacy report, the commitment of trader report (COT) with data dating back to 1962, the CFTC distinguishes "commercial" and "non-commercial" market participants where a "commercial" participant is defined as one "[...] engaged in business activities hedged by the use of futures and option markets"^[See [CFTC regulation 1.3, 17 CFR 1.3(z)](https://www.gpo.gov/fdsys/pkg/CFR-1998-title17-vol1/xml/CFR-1998-title17-vol1-sec1-3.xml) for details.]. A third category, "non-reportable", aggregates positions for participants not meeting the reporting threshold. In response to concerns related to category accuracy^[See for example @ederington_who_2002.] the CFTC now refines its classification in the disaggregated commitment of trader report (DCOT) where participant categories include "Producer/Merchant/Processor/User", "Swap dealer", "Money manager" and "Other reportable" as well as in the supplemental commitment of trader report (SCOT) that details commodity index trader aggregate positions for thirteen agricultural commodity futures markets. Data for DCOT and SCOT date back to 2006 and 2007 respectively.  
We examine the period from 1997 to 2018, during which we utilize futures-only position data from the COT report---the only CFTC report providing data for the entire duration of interest---as well as futures term structure price and open interest data sourced from Bloomberg. The sample period is further divided into four distinct intervals. The pre-financialisation era spans the 1997-2003 period and is naturally followed by the financialisation phase running from 2004 to 2008 which in turn leads into the financial crisis and its aftermath, covering 2008 to 2013. The final interval starts in 2013 with the onset of tapering and the conclusion of the US Quantitative Easing (QE) program.

@@

# Tables

@@

# Appendix

>>>>

The commodity futures trading commission (CFTC) operates a comprehensive system of collecting information on market participants known as the large trader reporting program (LTRP) where it records and reports to the public position data on market participants who have position levels in excess of a particular market specific threshold^[Current reporting level thresholds available at: [www.ecfr.gov](https://www.ecfr.gov/cgi-bin/retrieveECFR?gp=&SID=970471b8455f4bab7db4110cfde50731&mc=true&r=SECTION&n=se17.1.15_103).]. The commission collects market data and position information daily from clearing members, futures commission merchants (FCMs) and foreign brokers and publishes corresponding summaries weekly in a series of market specific reports. In these reports, individual traders are categorised according to the nature of their trading activity on the basis of self-reported information^[CFTC form 40. Available at: [www.cftc.gov](https://www.cftc.gov/sites/default/files/idc/groups/public/@forms/documents/file/cftcform40.pdf).] that is subject to review by CFTC staff for reasonableness. Position information is then aggregated by category and asset type with each report providing an aggregation for a set of report specific categories and, for each category, a breakdown between futures only positions and positions in futures and options combined.  
In its legacy report, the commitment of trader report (COT) with data dating back to 1962, the CFTC distinguishes "commercial" and "non-commercial" market participants where a "commercial" participant is defined as one "[...] engaged in business activities hedged by the use of futures and option markets"^[See [CFTC regulation 1.3, 17 CFR 1.3(z)](https://www.gpo.gov/fdsys/pkg/CFR-1998-title17-vol1/xml/CFR-1998-title17-vol1-sec1-3.xml) for details.]. A third category, "non-reportable", aggregates positions for participants not meeting the reporting threshold. In response to concerns related to category accuracy^[See for example @ederington_who_2002.] the CFTC now refines its classification in the disaggregated commitment of trader report (DCOT) where participant categories include "Producer/Merchant/Processor/User", "Swap dealer", "Money manager" and "Other reportable" as well as in the supplemental commitment of trader report (SCOT) that details commodity index trader aggregate positions for thirteen agricultural commodity futures markets. Data for DCOT and SCOT date back to 2006 and 2007 respectively.  
We examine the period from 1997 to 2018, during which we utilise futures-only position data from the COT report---the only CFTC report providing data for the entire duration of interest---as well as futures term structure price and open interest data sourced from Bloomberg. The sample period is further divided into four distinct intervals. The pre-financialisation era spans the 1997-2003 period and is naturally followed by the financialisation phase running from 2004 to 2008 which in turn leads into the financial crisis and its aftermath, covering 2008 to 2013. The final interval starts in 2013 with the onset of tapering and the conclusion of the US Quantitative Easing (QE) program.

@@

# Tables {.unnumbered}

@@

# Appendix {.unnumbered}

### 2025-04-04 docs/paper.tex
Recompile for resubmission

The commodity futures trading commission (CFTC) operates a comprehensive
system of collecting information on market participants known as the
large trader reporting program (LTRP) where it records and reports to
the public position data on market participants who have position levels
in excess of a particular market specific threshold\footnote{Current
  reporting level thresholds available at:
  \href{https://www.ecfr.gov/cgi-bin/retrieveECFR?gp=&SID=970471b8455f4bab7db4110cfde50731&mc=true&r=SECTION&n=se17.1.15_103}{www.ecfr.gov}.}.
The commission collects market data and position information daily from
clearing members, futures commission merchants (FCMs) and foreign
brokers and publishes corresponding summaries weekly in a series of
market specific reports. In these reports, individual traders are
categorised according to the nature of their trading activity on the
basis of self-reported information\footnote{CFTC form 40. Available at:
  \href{https://www.cftc.gov/sites/default/files/idc/groups/public/@forms/documents/file/cftcform40.pdf}{www.cftc.gov}.}
that is subject to review by CFTC staff for reasonableness. Position
information is then aggregated by category and asset type with each
report providing an aggregation for a set of report specific categories
and, for each category, a breakdown between futures only positions and
positions in futures and options combined.\\
In its legacy report, the commitment of trader report (COT) with data
dating back to 1962, the CFTC distinguishes ``commercial'' and
``non-commercial'' market participants where a ``commercial''
participant is defined as one ``{[}\ldots{]} engaged in business
activities hedged by the use of futures and option markets''\footnote{See
  \href{https://www.gpo.gov/fdsys/pkg/CFR-1998-title17-vol1/xml/CFR-1998-title17-vol1-sec1-3.xml}{CFTC
  regulation 1.3, 17 CFR 1.3(z)} for details.}. A third category,
``non-reportable'', aggregates positions for participants not meeting
the reporting threshold. In response to concerns related to category
accuracy\footnote{See for example \citet{ederington_who_2002}.} the CFTC
now refines its classification in the disaggregated commitment of trader
report (DCOT) where participant categories include
``Producer/Merchant/Processor/User'', ``Swap dealer'', ``Money manager''
and ``Other reportable'' as well as in the supplemental commitment of
trader report (SCOT) that details commodity index trader aggregate
positions for thirteen agricultural commodity futures markets. Data for
DCOT and SCOT date back to 2006 and 2007 respectively.\\
We examine the period from 1997 to 2018, during which we utilize
futures-only position data from the COT report---the only CFTC report
providing data for the entire duration of interest---as well as futures
term structure price and open interest data sourced from Bloomberg. The
sample period is further divided into four distinct intervals. The
pre-financialisation era spans the 1997-2003 period and is naturally
followed by the financialisation phase running from 2004 to 2008 which
in turn leads into the financial crisis and its aftermath, covering 2008
to 2013. The final interval starts in 2013 with the onset of tapering
and the conclusion of the US Quantitative Easing (QE) program.

@@

\section{Tables}\label{tables}

@@

\section{Appendix}\label{appendix}

>>>>

The commodity futures trading commission (CFTC) operates a comprehensive
system of collecting information on market participants known as the
large trader reporting program (LTRP) where it records and reports to
the public position data on market participants who have position levels
in excess of a particular market specific threshold\footnote{Current
  reporting level thresholds available at:
  \href{https://www.ecfr.gov/cgi-bin/retrieveECFR?gp=&SID=970471b8455f4bab7db4110cfde50731&mc=true&r=SECTION&n=se17.1.15_103}{www.ecfr.gov}.}.
The commission collects market data and position information daily from
clearing members, futures commission merchants (FCMs) and foreign
brokers and publishes corresponding summaries weekly in a series of
market specific reports. In these reports, individual traders are
categorised according to the nature of their trading activity on the
basis of self-reported information\footnote{CFTC form 40. Available at:
  \href{https://www.cftc.gov/sites/default/files/idc/groups/public/@forms/documents/file/cftcform40.pdf}{www.cftc.gov}.}
that is subject to review by CFTC staff for reasonableness. Position
information is then aggregated by category and asset type with each
report providing an aggregation for a set of report specific categories
and, for each category, a breakdown between futures only positions and
positions in futures and options combined.\\
In its legacy report, the commitment of trader report (COT) with data
dating back to 1962, the CFTC distinguishes ``commercial'' and
``non-commercial'' market participants where a ``commercial''
participant is defined as one ``{[}\ldots{]} engaged in business
activities hedged by the use of futures and option markets''\footnote{See
  \href{https://www.gpo.gov/fdsys/pkg/CFR-1998-title17-vol1/xml/CFR-1998-title17-vol1-sec1-3.xml}{CFTC
  regulation 1.3, 17 CFR 1.3(z)} for details.}. A third category,
``non-reportable'', aggregates positions for participants not meeting
the reporting threshold. In response to concerns related to category
accuracy\footnote{See for example \citet{ederington_who_2002}.} the CFTC
now refines its classification in the disaggregated commitment of trader
report (DCOT) where participant categories include
``Producer/Merchant/Processor/User'', ``Swap dealer'', ``Money manager''
and ``Other reportable'' as well as in the supplemental commitment of
trader report (SCOT) that details commodity index trader aggregate
positions for thirteen agricultural commodity futures markets. Data for
DCOT and SCOT date back to 2006 and 2007 respectively.\\
We examine the period from 1997 to 2018, during which we utilise
futures-only position data from the COT report---the only CFTC report
providing data for the entire duration of interest---as well as futures
term structure price and open interest data sourced from Bloomberg. The
sample period is further divided into four distinct intervals. The
pre-financialisation era spans the 1997-2003 period and is naturally
followed by the financialisation phase running from 2004 to 2008 which
in turn leads into the financial crisis and its aftermath, covering 2008
to 2013. The final interval starts in 2013 with the onset of tapering
and the conclusion of the US Quantitative Easing (QE) program.

@@

\section*{Tables}\label{tables}
\addcontentsline{toc}{section}{Tables}

@@

\section*{Appendix}\label{appendix}
\addcontentsline{toc}{section}{Appendix}
